Opportunities for improvement in the care of adult trauma patients arriving in shock

### 2024-11-27 manuscript.Rmd
ny plats

---
title: "Manuscript title here"
author: "Your name here"
output:
  html_document: default
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
---

>>>>

---
title: "Opportunities for improvement in the care of adult trauma patients arriving in shock"
author: "Carl Skog"
output:
  html_document: default
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
---

### 2024-11-27 thesis.Rmd
ny plats

---
title: "Opportunities for improvement in the care of adult trauma patients arriving in shock"
subtitle: "Add a Subtitle if Needed"
author: "Carl Skog"
output:
  html_document: default
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "out") })
editor_options: 
  markdown: 
    wrap: 72
---

@@

```{r sample-characteristics, echo=FALSE}
sample.characteristics.table
```

## Descriptive data
The study demographics median age is `r age` and the most common gender is male, `r male`. Clinically, the median ISS is (...) which is class (...), median INR is (...), and the most common Preinjury ASA class is (...), at (...)%.

```{r sample-characteristics BE, echo=FALSE}
sample.characteristics.tableBE
```

## Outcome data
Out of the `r sample` patients reviewed, `r ofi` patients had at least one positive OFI.

The most common

Table 6 and 7 present sub-ofi broad

```{r ofitable, echo=FALSE, results='asis'}
library(kableExtra)

ofi_table %>%
  kbl(
    col.names = c("BE Shock Class", "OFI Category", "Count", "Percentage (%)"),
    caption = "Counts and Percentages of OFI Broad Categories by BE Shock Class"
  ) %>%
  kable_classic(full_width = FALSE) %>%
  row_spec(0, bold = TRUE) %>% # Optional: Makes header bold
  group_rows(index = table(ofi_table$BE_class)) # Automatically merges BE_class rows

```

```{r combined_ofitable, echo=FALSE, results='asis'}
library(kableExtra)

combined_table %>%
  kbl(
    col.names = c(
      "BE Shock Class",
      "OFI Category",
      "BE Count",
      "BE Percentage (%)",
      "SBP Shock Class",
      "SBP Count",
      "SBP Percentage (%)"
    ),
    caption = "Counts and Percentages of OFI Broad Categories by BE and SBP Shock Classes"
  ) %>%
  kable_classic(full_width = FALSE) %>%
  row_spec(0, bold = TRUE) # Optional: Makes header bold

```

```{r BE-subofi, echo=FALSE}
print(p)
```

```{r SBP-subofi, echo=FALSE}
print(pSBP)
```

Table 8 and 9 present sub-ofi detailed

```{r SBP-subofi-detailed, echo=FALSE}
print(pd)
```

```{r BE-subofi-detailed, echo=FALSE}
print(pSBPd)
```

Table 10 and 11 adjusted log
```{r Adjusted_log, echo=FALSE}
log_regBE_sample.characteristics.table
log_regSBP_sample.characteristics.table
```

Table 12 and 13 unadjusted log
```{r Unadjusted_log, echo=FALSE}
log_regBE_sample.characteristics.table_unadjusted
log_regSBP_sample.characteristics.table_unadjusted
```

You can also embed plots:

```{r plot, echo=FALSE}
plot(pressure)
```

You can also mix text and code, so called inline code, like this: `r 2+5`.

>>>>

---
title: "Opportunities for improvement in the care of adult trauma patients arriving in shock"
subtitle: "Add a Subtitle if Needed"
author: "Carl Skog"
output:
  html_document: default
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = "/home/carls/out/thesis.html") })
editor_options: 
  markdown: 
    wrap: 72
---

@@

Table 3 - Sample characteristics showing missing data.
```{r sample-characteristics, echo=FALSE}
sample.characteristics.table
```

## Descriptive data
The study demographics median age is `r age` and the most common gender is male, `r male`. Clinically, the median ISS is 5 which is considered minor injuries, median INR is 1,00, and the most common Preinjury ASA class is 1, at 62%.

Table 4 - Sample characteristics used in regression.
```{r sample-characteristics BE, echo=FALSE}
sample.characteristics.table_reg
```

## Outcome data
Out of the `r log_reg_count` patients reviewed without missing data, `r ofi` patients had at least one positive OFI.

The most common broad sub-ofi category was clinical judgement error, at 103 (37%). The other three major broad sub-ofis are: inadequate resources 61 (22%), delay in treatment 49 (17%) and missed diagnosis 48 (17%). The smallest one, is other errors, at 5 (1,8%).

When classified, the most common broad sub-ofi in class 1 (no shock) is clinical judgement error both in the BE and SBP groups, while the other sub-ofi also follow a similar distribution between the groups. The largest sub-ofi continuous to be clinical judgment error for BE group in class 2, 3 and 4, at 34,6%; 31,2%; and 42,9% in this order. On other hand, the largest sub-ofi in SBP class 2 is delay in treatment at 50%. The SBP class 3 is evenly distributed between clinical judgement error and missed diagnosis at 40% each. Lastly, class 4 SBP's most common sub-ofi is clinical judgment error 46,7%.

Overall 3989 (81%) of the patients were classified as having no shock (class 1) according to the BE classification, the sum of the remaining classes accounted for 930 (19%) patients. The distribution of the classes were decreasing in numbers with increased severity, with class 2: 644 (13%), class 3: 170 (3,5%), and class 4: 116 (2,4%).

The SBP classification system sorted 4455 (91%) patients as having no shock (class 1), while the remaining classes with patients in shock accounted for 464 (9%) of all the patients. Among the classes defined with shock, class 2 was the biggest at 239 (4,9%), followed by class 4 at 127 (2,6%), and lastly class 3 with 99 (2,0%).

Table 5 - Sub-OFI, classified according to BE

```{r sub-ofi BE, echo=FALSE}
BEsubofi_tbl
```

Table 6 - Sub-OFI, classified according to SBP

```{r sub-ofi SBP, echo=FALSE}
SBPsubofi_tbl
```

## Main results
The adjusted analyses showed only statistical significance when comparing BE class 4 to class 1, which had 63% lower odds for OFI (OR 0,37; 95% CI 0,14-0,83; p-value 0.026).

The un-adjusted analysis showed when comparing to class 1 in BE group, class 2 (OR 1,61; 95% CI 1,17-2,20; p-value 0.003) and 3 (OR 1,91; 95% CI 1,08-3,16; p-value 0.018) were significant in increased risk for OFI. For the SBP group, comparison between class 1 and class 4 showed significant increased risk (OR 2,28; 95% CI 1,26-3,85; p-value 0.004).

Table 7 and 8 adjusted log
```{r Adjusted_log, echo=FALSE}
combined_tableBE
combined_tableSBP
```
